R Notebook

### 2020-04-24 Analysis/nat_stories.Rmd
changes from adding abstract graphs

---
title: "Natural Stories"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

@@

>>>>

---
title: "Natural Stories"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(viridis)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

@@

```
## Graphs for abstract
```{r error_abstract}
for_plot <- data_error_summ %>% 
  mutate(num_correct=case_when(
    critical_factor %in% c("1","2","3") ~ "3 or fewer",
    T ~ critical_factor)) %>% 
    mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct"))

ggplot(for_plot, aes(x=pct_correct, y=mean_rt, color=accurate))+
  geom_point(size=1)+
  scale_color_viridis(discrete=T)+
  labs(x="Fraction words selected correctly",
       y="Mean reaction time",
       color="?s correct")+
  coord_cartesian(xlim=c(0.4,1), ylim=c(0,1600))+
    scale_color_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
  guides(color=FALSE)

ggsave("../Papers/error.pdf", width=3, height=2, unit="in")

some <- data_filt %>% select(subject, critical) %>% 
  unique() %>% 
  left_join(data_error_summ, by=c("subject", "critical")) %>% 
  mutate(accurate=ifelse(pct_correct>.8,">80% correct", "<80% correct")) %>% 
  mutate(correct=critical*6) %>% 
  group_by(correct,accurate)

ggplot(some, aes(x=correct,fill=accurate))+
  geom_bar(position="dodge")+
  facet_grid(.~accurate)+
  labs(x="Comp. questions correct (out of 6)", y="Participants")+
  scale_fill_manual(values=c(">80% correct"="darkgreen","<80% correct"="darkred"))+
guides(fill=FALSE)

ggsave("../Papers/comp.pdf", width=3, height=1.5, unit="in")

### 2020-04-24 Analysis/nat_stories_lms.Rmd
changes from adding abstract graphs

---
title: "Natural Stories LM analysis"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(mgcv)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

@@

past_word <- labs %>% 
  mutate(word_num=word_num+1) %>% 
  rename(past_txl=txl_center, past_ngram=ngram_center, past_grnn=grnn_center, past_freq=freq, past_length=length) %>% 
  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
  left_join(labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))

labelled <- data_ready %>% left_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  select(rt, subject, Word_In_Story_Num, Story_Num,txl_center,ngram_center,grnn_center, freq_center,length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("../Data/for_modelling.Rds")

@@

```{r lm}

lm_ngram_model <- lm(rt ~ ngram_center + freq_center * length_center+ past_ngram + past_freq*past_length, data=labelled)

lm_txl_model <- lm(rt ~ txl_center + freq_center * length_center+ past_txl + past_freq*past_length, data=labelled)

lm_grnn_model <- lm(rt ~ grnn_center + freq_center * length_center+ past_grnn + past_freq*past_length, data=labelled)

```{r lm_summary}

summary(lm_ngram_model)

summary(lm_txl_model)

summary(lm_grnn_model)

```
<!--
# Now with heirarchy

@@

# show_summary <- function(model){
#   intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
#   stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)
#   
#   stats
# }

@@

gam_ngram <- gam(rt ~ s(ngram_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)

gam_txl <- gam(rt ~ s(txl_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)

gam_grnn <- gam(rt ~ s(grnn_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=filter(labelled))

@@

```{r}
 plot.gam(gam_ngram, select =1)
plot.gam(gam_ngram, select =3)

 plot.gam(gam_txl, select =1)
plot.gam(gam_txl, select =3)

 plot.gam(gam_grnn, select =1)
plot.gam(gam_grnn, select =3)

@@

```{r density}

ggplot(labelled, aes(x=ngram_center))+geom_density()

ggplot(labelled, aes(x=txl_center))+geom_density()

ggplot(labelled, aes(x=grnn_center))+geom_density()

@@

>>>>

---
title: "Natural Stories LM analysis"
output: github_document
---
```{r, include=F}
knitr::opts_chunk$set(echo = FALSE, warning=F)
options(knitr.table.format = "html")
library(tidyverse)
library(readr)
library(brms)
library(lme4)
library(rstan)
library(tidybayes)
library(knitr)
library(mgcv)
library(mgcViz)
theme_set(theme_bw())
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

@@

# past_word <- labs %>% 
#   mutate(word_num=word_num+1) %>% 
#   rename(past_txl=txl_center, past_ngram=ngram_center, past_grnn=grnn_center, past_freq=freq, past_length=length) %>% 
#   select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
#   left_join(labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))

# labelled <- data_ready %>% left_join(past_word, by=c("word_num", "word", "sentence")) %>%
#   filter(word_num>1) %>% 
#   select(rt, subject, Word_In_Story_Num, Story_Num,txl_center,ngram_center,grnn_center, freq_center,length_center, starts_with("past")) %>% 
#   mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
#   select(-Word_In_Story_Num, -Story_Num) %>% write_rds("../Data/for_modelling.Rds")

@@

<!--
# Now with heirarchy

@@

show_summary <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()
  stats <- gather_draws(model, `b_.*`, regex=T) %>% mutate(above_0=ifelse(.value>0, 1,0)) %>% group_by(.variable) %>% summarize(pct_above_0=mean(above_0)) %>% mutate(pval_equiv = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% left_join(intervals, by=".variable") %>% select(term=.variable, estimate=.value, lower=.lower, upper=.upper, pval_equiv)

  stats
}

@@

# gam_ngram <- gam(rt ~ s(ngram_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
# 
# gam_txl <- gam(rt ~ s(txl_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled)
# 
# gam_grnn <- gam(rt ~ s(grnn_center, bs="cr", k=20)+ te(freq_center, length_center, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=filter(labelled))

@@

```{r}
#  plot.gam(gam_ngram, select =1)
# plot.gam(gam_ngram, select =3)
# 
#  plot.gam(gam_txl, select =1)
# plot.gam(gam_txl, select =3)
# 
#  plot.gam(gam_grnn, select =1)
# plot.gam(gam_grnn, select =3)

@@

```{r density}
# 
# ggplot(labelled, aes(x=ngram_center))+geom_density()
# 
# ggplot(labelled, aes(x=txl_center))+geom_density()
# 
# ggplot(labelled, aes(x=grnn_center))+geom_density()

@@

# Second approach, being more selective
This time around we only include words that are single tokens in *all* the models and that have full data. We need the full data both for previous word and current word for the models.

We try both on pre-error only data and on all data (modulo exclusions). This is all uncentered.

```{r labs2}
# A more selective approach

select_labs <- labs %>% 
  filter(ngram_token_count==1 & txl_token_count==1 & grnn_token_count==1) %>% 
  filter(!is.na(txl_surp) & !is.na(ngram_surp) & !is.na(grnn_surp) & !is.na(freq) & !is.na(length)) %>% 
    mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T))

past_word <- select_labs %>% 
  mutate(word_num=word_num+1) %>% 
  rename(past_txl=txl_surp, past_ngram=ngram_surp, past_grnn=grnn_surp, past_freq=freq, past_length=length,past_c_txl=txl_center, past_c_ngram=ngram_center, past_c_grnn=grnn_center, past_c_freq=freq_center, past_c_length=length_center) %>% 
  select(sentence,Story_Num,Sentence_Num, word_num, starts_with("past")) %>% 
  left_join(select_labs, by=c("Story_Num","Sentence_Num","sentence", "word_num"))

data_pre_error <- data_no_first %>% filter(rt>100 & rt<5000) %>% 
  filter(after_mistake==0) %>% 
  select(subject, word_num, word, rt, sentence, type)

labelled_pre_error <- data_pre_error %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
  select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("pre_error.rds")

labelled_post_error <- data_ready %>% inner_join(past_word, by=c("word_num", "word", "sentence")) %>%
  filter(word_num>1) %>% 
select(rt, subject, Word_In_Story_Num, Story_Num,txl_surp,ngram_surp,grnn_surp, freq,length,
         txl_center, ngram_center,grnn_center,freq_center, length_center, starts_with("past")) %>% 
  mutate(Word_ID=as_factor(str_c(Story_Num, Word_In_Story_Num, sep="_"))) %>% 
  select(-Word_In_Story_Num, -Story_Num) %>% write_rds("post_error.rds")

## Linear models

Using pre-error data only.

```{r lm-2}

lm_ngram_model_1 <- lm(rt ~ ngram_surp + freq * length+ past_ngram + past_freq*past_length, data=labelled_pre_error)

lm_txl_model_1 <- lm(rt ~ txl_surp + freq * length+ past_txl + past_freq*past_length, data=labelled_pre_error)

lm_grnn_model_1 <- lm(rt ~ grnn_surp + freq * length + past_grnn + past_freq*past_length, data=labelled_pre_error)

```{r lm_summary-2}

summary(lm_ngram_model_1)

summary(lm_txl_model_1)

summary(lm_grnn_model_1)

```
Including post-error data.

```{r lm-3}

lm_ngram_model_2 <- lm(rt ~ ngram_surp + freq * length+ past_ngram + past_freq*past_length, data=labelled_post_error)

lm_txl_model_2 <- lm(rt ~ txl_surp + freq * length+ past_txl + past_freq*past_length, data=labelled_post_error)

lm_grnn_model_2 <- lm(rt ~ grnn_surp + freq * length + past_grnn + past_freq*past_length, data=labelled_post_error)

```{r lm_summary-3}

summary(lm_ngram_model_2)

summary(lm_txl_model_2)

summary(lm_grnn_model_2)

## GAMs

Using only pre-error data.

```{r gams-2}
gam_ngram_1 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)

gam_txl_1 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)

gam_grnn_1 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_pre_error)

```{r plot_gam-2}
 plot.gam(gam_ngram_1, select =1)
plot.gam(gam_ngram_1, select =3)

 plot.gam(gam_txl_1, select =1)
plot.gam(gam_txl_1, select =3)

 plot.gam(gam_grnn_1, select =1)
plot.gam(gam_grnn_1, select =3)

Including post-error data.

```{r gams-3}
gam_ngram_2 <- gam(rt ~ s(ngram_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_ngram, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)

gam_txl_2 <- gam(rt ~ s(txl_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_txl, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)

gam_grnn_2 <- gam(rt ~ s(grnn_surp, bs="cr", k=20)+ te(freq, length, bs="cr")+s(past_grnn, bs="cr", k=20)+te(past_freq, past_length, bs="cr"), data=labelled_post_error)

```{r plot_gam-3}
 plot.gam(gam_ngram_2, select =1)
plot.gam(gam_ngram_2, select =3)

plot_smooths(model=gam_ngram_2,series=ngram_surp)
plot_smooths(model=gam_ngram_2,series=past_ngram)

 plot.gam(gam_txl_2, select =1)
plot.gam(gam_txl_2, select =3)

plot_smooths(model=gam_txl_2,series=txl_surp)
plot_smooths(model=gam_txl_2,series=past_txl)

 plot.gam(gam_grnn_2, select =1)
plot.gam(gam_grnn_2, select =3)

 v <- seq(0,25, by=1)
a <- predict(gam_ngram_2, type="terms")

b <- predict_gam(gam_ngram_2, exclude_terms="te(freq,length)",length_out=2)

summary(gam_ngram_2)
```

This is much more reassuring. We do want to check with brms though.

priors <- c(
  set_prior("normal(1000, 1000)", class="Intercept"),
  set_prior("normal(0, 500)", class="b"),
  set_prior("normal(0, 500)", class="sd"),
  set_prior("lkj(1)",       class="cor"))

brm_ngram_model <- brm(rt ~ ngram_surp + freq * length + past_ngram + past_freq*past_length + 
                         (ngram_surp + freq * length + past_ngram + past_freq*past_length|subject), data=labelled_pre_error, file="brm_ngram_pre", prior=priors)

summary(brm_ngram_model)

show_summary(brm_ngram_model)
```

### 2020-04-24 Prep_code/nat_stories_prep.Rmd
changes from adding abstract graphs

```{r conglomerate}
all <- txl_translator %>% 
  left_join(ngram_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
  left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  filter(Word_In_Sentence_Num!=0) %>% 
  ungroup() %>% 
  mutate(txl_center=txl_surp-mean(txl_surp, na.rm=T),
         ngram_center=ngram_surp-mean(ngram_surp, na.rm=T),
         grnn_center=grnn_surp-mean(grnn_surp, na.rm=T),
         freq_center=freq-mean(freq, na.rm=T),
         length_center=length-mean(length, na.rm=T)) %>% 
  write_rds("natural_stories_surprisals.rds")

>>>>

```{r conglomerate}
all <- txl_translator %>% 
  left_join(ngram_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  left_join(grnn_translator, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>%
  left_join(freq, by=c("Story_Num", "Sentence_Num","Word_In_Sentence_Num", "Word", "Word_In_Story_Num")) %>% 
  filter(Word_In_Sentence_Num!=0) %>% 
  ungroup() %>% 
  write_rds("natural_stories_surprisals.rds")
